---
title: TidyTuesday - School Diversity
author: Kevin Putschko
date: '2019-09-26'
slug: tidytuesday-school-diversity
categories:
  - R
tags:
  - TidyTuesday
---

```{r setup}
pacman::p_load(tidyverse, colorspace)

data <- 
  read_csv("https://raw.githubusercontent.com/rfordatascience/tidytuesday/master/data/2019/2019-09-24/school_diversity.csv")
```

Hello!

Welcome back! It's been a while since I've written a blog post. So let's get started, shall we.

### Introduction

This data comes from the Tidy Tuesday website, dealing with school diversity across the USA.  Let's begin by taking a look a the data.  The data pertains to two school years, the 1994-95 year and 2016-17 school year.  Perhaps we can do a simple comparison of the two years for Michigan.

```{r}
data %>% glimpse()
data %>% count(SCHOOL_YEAR)
```

### Research Question

How has school diversity changed in Michigan between 1994 and 2016?

```{r}
data %>% 
  filter(ST == "MI") %>% 
  separate(d_Locale_Txt, into = c("Location", "Designation")) %>% 
  pivot_longer(cols = AIAN:Multi, names_to = "Ethnicity", values_to = "Proportion") %>% 
  group_by(SCHOOL_YEAR, Location, Ethnicity) %>% 
  summarise(Proportion = mean(Proportion, na.rm = TRUE)) %>%
  pivot_wider(names_from = SCHOOL_YEAR, values_from = Proportion) %>% 
  filter(!is.na(Location)) %>% 
  
  ggplot() +
  aes(x = Ethnicity) +
  geom_segment(aes(y = `1994-1995`, yend = `2016-2017`, 
                   x = Ethnicity, xend = Ethnicity), 
               size = 3.5, 
               color = "grey80") +
  geom_point(aes(y = `1994-1995`), 
             fill = "#fc8d59", color = darken("#fc8d59", 0.30), 
             size = 3, shape = 21) +
  geom_point(aes(y = `2016-2017`),
             fill = "#99d594", color = darken("#99d594", 0.30), 
             size = 3, shape = 21) +
  geom_text(data = . %>% filter(Ethnicity == "White", Location == "city"), 
            mapping = aes(x = Ethnicity, y = `1994-1995`),
            label = "1994/95", 
            color = "#fc8d59",
            fontface = "bold",
            vjust = 2) +
  facet_wrap("Location", labeller = labeller(Location = str_to_title)) +
  geom_text(data = . %>% filter(Ethnicity == "White", Location == "city"), 
            mapping = aes(x = Ethnicity, y = `2016-2017`),
            label = "2016/17", 
            color = "#99d594",
            fontface = "bold",
            vjust = 2) +
  facet_wrap("Location", labeller = labeller(Location = str_to_title)) +
  coord_flip() +
  theme_minimal() +
  labs(x = NULL, y = "Proportion", 
       title = "Ethnic Diversity in Michigan Schools", subtitle = "By Location",
       caption = "Source: National Center for Education Statistics\nGraph adopted from @Mathl3t3")
```

### Conclusion

Time is short this week, so I'll leave it there for now.  I'll be back with more next time! As always the code will be available on my [GitHub](https://github.com/kputschko/kp-blogdown/tree/master/content/post) page.  

