---
title: Seattle Pet Names
author: Kevin Putschko
date: '2019-03-26'
slug: seattle-pet-names
categories:
  - Data Viz
tags:
  - TidyTuesday
---

## Introduction

```{r setup}
knitr::opts_chunk$set(echo = TRUE)
pacman::p_load(tidyverse, kp.helpers, lubridate, sf, zipcode, leaflet)

data_raw <- 
  read_csv("https://raw.githubusercontent.com/rfordatascience/tidytuesday/master/data/2019/2019-03-26/seattle_pets.csv")
```

This week I'll be exploring the names of pets registered in Seattle, WA.  The data comes from the [TidyTuesday](https://github.com/rfordatascience/tidytuesday/tree/master/data/2019/2019-03-26) weekly blog. 

We see there are 7 columns and 52,519 rows, with columns relating to data, name, species, breed, and zip code.  

```{r echo=TRUE}
data_raw %>% glimpse()
```

The first thing I notice is that the date is provided in character format.  Let's convert that to a date format instead.

```{r dates, echo=TRUE}
data_dates <- 
  data_raw %>% 
  mutate(license_issue_date = as.Date(license_issue_date, format = "%B %d %Y"))
```

Now let's take a look at a summary table.

```{r summary, echo=TRUE}
data_dates %>% fx_describe() %>% DT::datatable()
```

Alright! We see the dates range from 2003 to 2018, and that there are some missing values for secondary breed, name and zip code.

With this in mind, let's take a look at 1) Name Popularity, and 2) Species Popularity by Zip Code.

## Name Popularity 

First of all, let's calculate the date.  Once that's done we see that the distirbution of names by date is heavily skewed to recent years.  This will adjust our plans.  

```{r pet_names, echo=TRUE}
data_dates %>% 
  filter(!is.na(animals_name)) %>% 
  mutate(year = year(license_issue_date)) %>% 
  count(year) %>% 
  ggplot(aes(x = factor(year), y = n)) +
  geom_col() +
  scale_y_log10() +
  theme_kp() +
  labs(x = "Year", y = "Count (Log10)", title = "Records per Year")
```

So, let's just dig into the top names!  

```{r name_plot, echo=TRUE}
data_name_plot <- 
  data_dates %>% 
  count(species, animals_name, sort = TRUE) %>% 
  filter(n > 5) %>%
  group_by(species) %>% 
  top_n(10, n) %>%
  replace_na(list(animals_name = "No Recorded Name")) %>% 
  group_by(animals_name) %>% 
  mutate(total_names = sum(n, na.rm = TRUE),
         p = scales::percent(n / total_names, accuracy = 1),
         label_coord = cumsum(n) - n/2)
  
data_name_plot %>% 
  ggplot(aes(x = reorder(animals_name, total_names), y = n, fill = species)) +
  geom_col(position = "stack", color = "black") +
  geom_text(aes(y = label_coord, label = p), color = "white", size = 3, family = "sans") +
  coord_flip() +
  theme_kp() +
  labs(x = NULL, y = "Count", fill = "Species", title = "Top 10 Names of Cats and Dogs")
```

We can see there are a handful of names that are used for both cats and dogs.  Interestingly, only cats have records without any given names.  I imagine this is due to catch and release programs that might exist in the Seattle area. 

## Species by Region

Now let's look at species by region.  I see that there are some registered pigs and goats, let's see where those are coming from while we're at it.

```{r species, echo=TRUE}
data_raw %>% 
  count(species, sort = TRUE) %>% 
  ggplot(aes(x = as_factor(species), y = n)) + 
  geom_col() + 
  scale_y_log10() + 
  theme_kp() + 
  labs(x = "Species", y = "Count (Log 10)", title = "Count of Species in Seattle Area")
```

This is actually my first time using leaflet, and only one of a handful of times that I've worked with data viz and maps.  Let's see how it goes! 

```{r zips, echo=TRUE}
data(zipcode)

data_species <- 
  data_raw %>% 
  mutate_at(vars(species), as_factor) %>% 
  count(zip_code, species, sort = TRUE) %>%   
  inner_join(zipcode, by = c("zip_code" = "zip"))

pal <- 
  colorFactor(
    palette = "Dark2", 
    domain = data_species$species %>% unique())

data_species %>% 
  leaflet() %>% 
  setView(lng = -122.3, lat = 47.6, zoom = 11) %>% 
  addProviderTiles(providers$Wikimedia) %>% 
  addCircleMarkers(lng = ~longitude, 
                   lat = ~latitude, 
                   color = ~pal(species),
                   radius = ~n*0.02,
                   weight = 1)
  

```

This was much more challenging than expected.  

## Conclusion

What have I learned?  Maps are much more challenging to work with than I thought they would be.  I hit my three hour limit while trying to figure out how to use leaflet, but it was good practice.  Next time I'll try to add a legend so we know which color represents which species.

That's all for now.  As always you can find the code on my [GitHub](https://github.com/kputschko/kp-blogdown/tree/master/content/post) page.
